Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-12 shanghai.tex
Minor changes to notation in the section using the snake lemma.

\newcommand{\cdef}[1]{{{\color{blue}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\GCb}{{\GC_0}}
\newcommand{\GCf}{{\GC_f}}
\newcommand{\GCiso}[1]{\GC(#1)_{/\textit{iso}}}
\newcommand{\GCbiso}[1]{\GCb(#1)_{/\text{iso}}}
\newcommand{\GCfiso}[1]{\GCf(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\Loc}{{\operatorname{Loc}}}
\newcommand{\trFrob}[1]{t_{#1}}

@@

%\input{snake}
\begin{theorem}\label{thm:snake}
  If $G$ is a smooth abelian group scheme, locally of finite type over
  $\Fq$, and the abelian group $\pi_0(G)(\bFq)$ (the geometric
  component group) is finitely generated, then the trace of Frobenius
  homomorphism
  \[
  \trFrob{G} : \GCiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx),
  \]
  from the Picard group of the category of geometrized characters of $G$ to
  the group of characters of $G(\Fq)$, is an isomorphism.
\end{theorem}

\begin{proof}
  Consider the short exact sequence in the category of smooth group
  schemes, locally of finite type over $\Fq$, defining the component
  group scheme for $G$:
  \begin{equation}\label{eq:pi0}
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \end{equation}
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker t_{\pi_0(G)} \dar & \arrow[dashed]{d} \ker \trFrob{G} & \ker t_{G^0} \dar & \\
    0 \rar & \GCiso{\pi_0(G)} \rar \dar{t_{\pi_0(G)}}
    & \GCiso{G} \rar \dar{\trFrob{G}} & \GCiso{G^0} \rar \dar{t_{G^0}} & 0\\
    0 \rar & \Hom(\pi_0(G)(\Fq),\EEx) \rar \dar
    & \arrow[dashed]{d} \Hom(G(\Fq),\EEx) \rar & \Hom(G^0(\Fq),\EEx) \rar \dar & 0\\
    & \coker t_{\pi_0(G)} & \coker \trFrob{G} &  \coker t_{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  By
  Proposition~\ref{prop:connected}, $\ker t_{G^0} =0$ and $\coker t_{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker t_{\pi_0(G)}=0$
  and $\coker t_{\pi_0(G)}=0$.

@@

  The sequence of Picard groups is right-exact: it is exact at
  $\GCiso{G^0}$ by Proposition~\ref{prop:restriction}; and
  it is exact at $\GCiso{G}$ because the left-square
  commutes (by Proposition~\ref{prop:pullback}), the bottom sequence
  is exact at $\Hom(G(\Fq),\EEx)$ and $\coker t_{\pi_0(G)} =0$
  (by Propositions~\ref{prop:etale} and \ref{prop:X}) and
  \[
  \image(\GCiso{\pi_0(G)} \to \GCiso{G}) \subseteq \ker(\GCiso{G} \to \GCiso{G^0}),
  \]
  by the last part of Proposition~\ref{prop:pullback}. Specifically,
  suppose $\cs{L}$ is a geometrized character of $G$ that is trivial on
  $G^0$. Then $\trFrob{\cs{L}}\vert_{G^0(\Fq)} = \trFrob{\cs{L}^0} =0$,
  by Proposition~\ref{prop:pullback} again. Since the sequence of
  character groups is exact at $\Hom(G(\Fq),\EEx)$, there is
  some $\chi \in \Hom(\pi_0(G)(\Fq),\EEx)$ such that the
  pull-back of $\chi$ along $G(\Fq)\to \pi_0(G)(\Fq)$ is
  $\trFrob{\cs{L}}$. By Propositions~\ref{prop:etale} and \ref{prop:X},
  there is a a geometrized character $\cs{L}_\chi$ on $\pi_0(G)$, unique up
  to isomorphism, such that $\trFrob{\cs{L}_\chi} = \chi$. By
  Proposition~\ref{prop:pullback} again, $\cs{L}$ is isomorphic to the
  image of $\cs{L}_\chi$ under the monoidal functor $\GC(\pi_0(G)) \to \GC(G)$.

>>>>

\newcommand{\cdef}[1]{{{\color{blue}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\GCb}{{\GC_0}}
\newcommand{\GCf}{{\GC_f}}
\newcommand{\GCiso}[1]{\GC(#1)_{/\textit{iso}}}
\newcommand{\GCbiso}[1]{\GCb(#1)_{/\text{iso}}}
\newcommand{\GCfiso}[1]{\GCf(#1)_{/\text{iso}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\Loc}{{\operatorname{Loc}}}
\newcommand{\trFrob}[1]{t_{#1}}

@@

%\input{snake}
\begin{theorem}\label{thm:snake}
  If $G$ is a smooth abelian group scheme, locally of finite type over
  $\Fq$, and the abelian group $\pi_0(G)(\bFq)$ (the geometric
  component group) is finitely generated, then the trace of Frobenius
  homomorphism
  \[
  \trFrob{G} : \GCiso{G} \to G(\Fq)^*
  \]
  from the Picard group of the category of geometrized characters of $G$ to
  the group of characters of $G(\Fq)$, is an isomorphism.
\end{theorem}

\begin{proof}
  Consider the short exact sequence in the category of smooth group
  schemes, locally of finite type over $\Fq$, defining the component
  group scheme for $G$:
  \begin{equation}\label{eq:pi0}
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \end{equation}
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker t_{\pi_0(G)} \dar & \arrow[dashed]{d} \ker \trFrob{G} & \ker t_{G^0} \dar & \\
    0 \rar & \GCiso{\pi_0(G)} \rar \dar{t_{\pi_0(G)}}
    & \GCiso{G} \rar \dar{\trFrob{G}} & \GCiso{G^0} \rar \dar{t_{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow[dashed]{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
    & \coker t_{\pi_0(G)} & \coker \trFrob{G} &  \coker t_{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  By
  Proposition~\ref{prop:connected}, $\ker t_{G^0} =0$ and $\coker t_{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker t_{\pi_0(G)}=0$
  and $\coker t_{\pi_0(G)}=0$.

@@

  The sequence of Picard groups is right-exact: it is exact at
  $\GCiso{G^0}$ by Proposition~\ref{prop:restriction}; and
  it is exact at $\GCiso{G}$ because the left-square
  commutes (by Proposition~\ref{prop:pullback}), the bottom sequence
  is exact at $G(\Fq)^*$ and $\coker t_{\pi_0(G)} =0$
  (by Propositions~\ref{prop:etale} and \ref{prop:X}) and
  \[
  \image(\GCiso{\pi_0(G)} \to \GCiso{G}) \subseteq \ker(\GCiso{G} \to \GCiso{G^0}),
  \]
  by the last part of Proposition~\ref{prop:pullback}. Specifically,
  suppose $\cs{L}$ is a geometrized character of $G$ that is trivial on
  $G^0$. Then $\trFrob{\cs{L}}\vert_{G^0(\Fq)} = \trFrob{\cs{L}^0} =0$,
  by Proposition~\ref{prop:pullback} again. Since the sequence of
  character groups is exact at $G(\Fq)^*$, there is
  some $\chi \in \pi_0(G)(\Fq)^*$ such that the
  pull-back of $\chi$ along $G(\Fq)\to \pi_0(G)(\Fq)$ is
  $\trFrob{\cs{L}}$. By Propositions~\ref{prop:etale} and \ref{prop:X},
  there is a a geometrized character $\cs{L}_\chi$ on $\pi_0(G)$, unique up
  to isomorphism, such that $\trFrob{\cs{L}_\chi} = \chi$. By
  Proposition~\ref{prop:pullback} again, $\cs{L}$ is isomorphic to the
  image of $\cs{L}_\chi$ under the monoidal functor $\GC(\pi_0(G)) \to \GC(G)$.
